---
title: "Character Recognition with Logistic Regression"
author: "Parthasarathi Samantaray"
date: "21/01/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Character Recognition with Logistic Regression
I have loaded the tidyverse package to use the functionalities as and when required.

```{r}
library(tidyverse)
```


I have set the working directory to the folder with the 2 datasets. After looking at the data sets, I found that there is no heading in the CSV files. So, I have used the read.csv function with header = FALSE to load the datasets and saved as variables named train and test respectively.  

```{r}
train <- read.csv("mnist_train.csv", header = FALSE)
test <- read.csv("mnist_test.csv", header = FALSE)
```
Checking the dimension and variables names using the functions dim(), head() and colnames(). The variables are named from V1 till V785 with V1 containing the digit labels. 
```{r}
head(train,1)
```
There are 60000 rows in training data set and 10000 rows in the test data set.

```{r}
#str(train)
```
```{r}
train%>% dim()
test%>% dim() 

test %>% colnames()
```
I have renamed the column V1 as Label for better readability and tested the update using the colnames() function.
```{r}
# Rename a columns in Train and test 

colnames(train)[colnames(train)=="V1"] <- "Label"
colnames(test)[colnames(test)=="V1"] <- "Label"

test %>% colnames()

```
I have used the 2 provided function definitions rotate and plot_matrix to view one of the rows in the training dataset.

```{r}
rotate <- function(x) {
  return(t(apply(x, 2, rev)))
}
plot_matrix <- function(vec) {
  q <- matrix(vec, 28, 28, byrow = TRUE)
  nq <- apply(q, 2, as.numeric)
  image(rotate(nq), col = gray((0:255)/255))
}

```

```{r}
train[34,-1] %>% plot_matrix()
```
I checked the current distribution of digits in both the 2 data sets using table() function. 
```{r}
table(train$Label)
table(test$Label)

```
I renamed the labels as 1 if the digit is 9 or 0 otherwise. I checked the distribution of 0's and 1's as per earlier distribution and both the distributions are consistent. 

```{r}

train$Label = ifelse(train$Label==9 ,1,0)
test$Label = ifelse(test$Label==9 ,1,0)

table(train$Label)
table(test$Label) 
```
I trained the Logistic Regression Model with formula Label = F(Pixel1, Pixel2, Pixel3, ., Pixel 784) and as there are 2 output values, used family as binomial with the logit link. I have saved the model as charRecog.glm. 

```{r}

charRecog.glm <- glm(Label ~ . ,data = train, family = binomial())

```
From the summary, I found there 72 pixels/variables are don't add any information to the model. These variables have coefficients as NA and the summary comes with the warning "72 not defined because of singularities".

```{r}
summary(charRecog.glm)
```
Using the above model, I have predicted the digits with inputs from test dataset and saved as testPredict. 

```{r}
testPredict<-predict(charRecog.glm,newdata = test[,2:785],type = "response")
```

```{r}
summary(testPredict)
```
By assuming the probability higher than 0.5  is predicted as digit 9(1 in the current model), the confusion matrix of this model as follows:- 

```{r}
table(test$Label,testPredict>.5)

```
```{r}
print(paste("The Sensitivity of the model is : ",8799/(8799+233)))
print(paste("The Specificity of the model is : ",776/(766+192)))

```
